Introduccion to the Tidyverse

```{r}
install.packages("gapminder") 
install.packages("dplyr")
library(dplyr)
library(gapminder)
```

Filtering by year

```{r}
gapminder %>%
  filter(year==1957, country=="China")
  
```

Arranging observations by life expectancy
```{r}

gapminder %>%
  arrange(lifeExp)

# Sort in descending order of 
gpaminder %>%
  arrange(desc(lifeExp))

```

Filtering and arranging
```{r}
gapminder %>%
  filter(year ==1957) %>%
  arrange(desc(pop))
```

Using mutate to change or create a column
```{r}
gapminder %>%
  mutate(lifeExp=12*lifeExp)

#Use mutate to create a new collum called lifeExpMonths
gapminder %>%
  mutate(lifeExpMonths = 12 * lifeExp)
```

Combining filter, mutate and arrange
```{r}
gapminder %>%
  filter(year == 2007) %>%
  mutate(lifeExpMonths = 12 * lifeExp) %>%
  arrange(desc(lifeExpMonths))
```

Visualizing with ggplot2
```{r}
install.packages("ggplot2")
library(ggplot2)
gapminder_1952 <- gapminder %>% 
  filter(year==1952)

ggplot(gapminder_1952, aes(x=pop, y=gdpPercap))+
  geom_point()
```
```{r}
ggplot(gapminder_1952, aes(x=pop, y=lifeExp))+
  geom_point()
```

Puttin the x-axis on a log scale
```{r}

gapminder_1952 <- gapminder %>%
  filter(year==1952)

ggplot(gapminder_1952, aes(x=pop, y = lifeExp))+
  geom_point() +
  scale_x_log10()

```

Adding color to a scatter plot
```{r}

gapminder
ggplot(gapminder_1952, aes(x=pop, y = lifeExp, color = continent))+
  geom_point() +
  scale_x_log10()

```

Adding size and color to a plot
```{r}
ggplot(gapminder_1952, aes(x=pop, y = lifeExp, color = continent, size = gdpPercap))+
  geom_point() +
  scale_x_log10()
```

Faceting
Creating a subgraph for each continent

```{r}

ggplot(gapminder, aes(x=pop, y = lifeExp, color= continent, size=pop))+
  geom_point() +
  scale_x_log10()+
  facet_wrap(~ year)
```

